Carrier phase and amplitude estimation of $M$-ary phase shift keying using pilots and data

### 2012-05-13 paper.tex
Added to paper.

\begin{IEEEproof} (Theorem~\ref{thm:normality})
The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at half integers.  This precludes the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.  However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the maximiser $\hat{\lambda}_L$, so
\begin{align*}
G_L^\prime(\hat{\lambda}_L) &= 0 \\
&= \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
\end{align*}
Define the function $Q(\lambda) = \prob G_L^\prime(\lambda) - \prob G_L^\prime(0)$ and write
\begin{align*}
0 &= G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) + Q(\hat{\lambda}_L) \\
&= \sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) + \sqrt{L}Q(\hat{\lambda}_L).
\end{align*}
Lemma~\ref{lem:Qconv} shows that
\[
\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
\]
where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$.  Lemma~\ref{lem:empiricprocc} shows that
\[
\sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} G_L^\prime(0).
\]
It follows that
\[
0 = o_P(1) + \sqrt{L}G_L^\prime(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
\]
and rearranging gives,
\[
\sqrt{L} \hat{\lambda}_L = o_P(1) + \frac{\sqrt{L}G_L'(0)}{p + Hd  + o_P(1)}.
\]
Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance
\[
p \prob R_1^2 \sin^2(\Phi_1) + d \prob R_1^2 \sin^2\fracpart{\Phi_1}
\]
and it follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
\[
\frac{p \prob R_1^2 \sin^2(\Phi_1) + d \prob R_1^2 \sin^2\fracpart{\Phi_1}}{?}
\]
as required.

Define the function $H_L(\lambda) = \prob G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ} we have
\begin{align*}
\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L) + H_L(\hat{\lambda}_L) - G(0) \big)
\end{align*}
Lemma~\ref{lem:empiricprocforrho} shows that 
\[
\sqrt{L}\big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L 
\]
where $X_L = \sqrt{L}\big( G_L(0) - H_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
\[
\sqrt{L}\big( H_L(\hat{\lambda}_L) - G(0) \big) = \hat{\lambda}_L \big( ? + o_P(1) \big).
\]
It follows that
\[
\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
\]
Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and covariance
\[
?
\]
and Lemma~\ref{lem:XLcovlambdaL} shows that the covariance between $X_L$ and $\sqrt{L}\hat{\lambda}_L$ goes to zero as $L\rightarrow\infty$.  So, the distribution of $\sqrt{L}m_L$ converges to the normal with zero mean and variance
\[
\var X_L + \var \sqrt{L} \hat{\lambda}_L ?^2 = BLERG.
\]
The covariance between $\sqrt{L}m_L$ and $\sqrt{L}\hat{\lambda}_L$ is correspondinly
\[
?
\]
\end{IEEEproof}

@@

\begin{lemma}\label{lem:Qconv}
\[
\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
\]
where 
\[
H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
\]
\end{lemma}
\begin{IEEEproof}
Write
\[
Q(\lambda) = \prob G_L^\prime(\lambda) - \prob G_L^\prime(0) = \frac{\abs{P}}{L} k_1(\lambda) + \frac{\abs{D}}{L} k_2(\lambda)
\]
where
\begin{align}
k_1(\lambda) &= \prob R_1\big( \sin(\lambda + \Phi_1) - \sin(\Phi_1) \big), \;\;\; \text{and} \nonumber\\
k_2(\lambda) &= \prob R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big). \label{eq:k2def}
\end{align}
Since,
\[
\sin(\lambda + \phi) - \sin(\phi) = \lambda \big( \cos(\phi) + \zeta(\lambda) \big),
\]
where $\zeta(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero, then
\begin{align*}
k_1(\lambda) &= \prob R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
&= \lambda \big( \prob R_1\cos(\Phi_1) \; + \; \zeta(\lambda) \prob R_1 \big) \\
&= \lambda \big( 1 \; + \; \zeta(\lambda) K \big),
\end{align*}
since $h_1(0) = \prob R_1\cos(\Phi_1) = 1$ and where we put $K = \prob R_1$, a finite constant.  Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e. $\zeta(\hat{\lambda}_L) = o_P(1)$.  So, 
\[
k_1(\hat{\lambda}_L) = \hat{\lambda}_L \big( 1 \; + \; \zeta(\hat{\lambda}_L) K \big) = \hat{\lambda}_L \big( 1  + o_P(1)\big).
\]
Lemma~\ref{lem:k2} shows that
\[
k_2(\hat{\lambda}_L) = \hat{\lambda}_L \big( ? \; + \; o_P(1) \big).
\]
\end{IEEEproof}

\begin{lemma}\label{lem:k2}
With the function $k_2(\lambda)$ defined as in~\eqref{eq:k2def}, then
k_2(\hat{\lambda}_L) = \hat{\lambda}_L \big( H \; + \; o_P(1) \big).
H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
\end{lemma}
\begin{IEEEproof}
Now, if $0 \leq \lambda < \tfrac{\pi}{M}$, then for each integer $k$,

@@

&= \lambda\big(\cos\fracpart{\phi} + \zeta_k(\lambda)\big)

@@

when $\phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda )$ and 

@@

when $\phi \in [ \tfrac{(2k+1)\pi}{M} - \lambda,\tfrac{(2k+1)\pi}{M} )$, and where

@@

Observe that $v_k(\tfrac{(2k+1)\pi}{M}) = 2\sin(\tfrac{\pi}{M})$.  We can now write

@@

and since $v_k(\lambda,\phi) = 0$ for $\phi \in [0,2\pi)$ when $k \notin \{0,1,\dots,M-1\}$ the infinite sum above can be truncated so that

>>>>

\begin{IEEEproof} (Theorem~\ref{thm:normality})
The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at half integers.  This precludes the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.  However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the maximiser $\hat{\lambda}_L$, so
\begin{align*}
G_L^\prime(\hat{\lambda}_L) &= 0 \\
&= \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
\end{align*}
Define the function $Q(\lambda) = \prob G_L^\prime(\lambda) - \prob G_L^\prime(0)$ and write
\begin{align*}
0 &= G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) + Q(\hat{\lambda}_L) \\
&= \sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) + \sqrt{L}Q(\hat{\lambda}_L).
\end{align*}
Lemma~\ref{lem:Qconv} shows that
\[
\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
\]
where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$.  Lemma~\ref{lem:empiricprocc} shows that
\[
\sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} G_L^\prime(0).
\]
It follows that
\[
0 = o_P(1) + \sqrt{L}G_L^\prime(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
\]
and rearranging gives,
\[
\sqrt{L} \hat{\lambda}_L = o_P(1) + \frac{\sqrt{L}G_L'(0)}{p + Hd  + o_P(1)}.
\]
Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $A_1 + A_2$ and it follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
\[
\frac{A_1 + A_2}{(p + Hd)^2}.
\]

We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Define the function $H_L(\lambda) = \prob G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ} we have
\begin{align*}
\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L) + H_L(\hat{\lambda}_L) - G(0) \big)
\end{align*}
Lemma~\ref{lem:empiricprocforrho} shows that 
\[
\sqrt{L}\big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L 
\]
where $X_L = \sqrt{L}\big( G_L(0) - H_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
\[
\sqrt{L}\big( H_L(\hat{\lambda}_L) - G(0) \big) = \hat{\lambda}_L \big( ? + o_P(1) \big).
\]
It follows that
\[
\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
\]
Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and covariance
\[
?
\]
and Lemma~\ref{lem:XLcovlambdaL} shows that the covariance between $X_L$ and $\sqrt{L}\hat{\lambda}_L$ goes to zero as $L\rightarrow\infty$.  So, the distribution of $\sqrt{L}m_L$ converges to the normal with zero mean and variance
\[
\var X_L + \var \sqrt{L} \hat{\lambda}_L ?^2 = BLERG.
\]
The covariance between $\sqrt{L}m_L$ and $\sqrt{L}\hat{\lambda}_L$ is correspondinly
\[
?
\]
\end{IEEEproof}

@@

\begin{lemma}\label{lem:Qconv}
\[
\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
\]
where 
\[
H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
\]
\end{lemma}
\begin{IEEEproof}
Write
\[
Q(\lambda) = \prob G_L^\prime(\lambda) - \prob G_L^\prime(0) = \frac{\abs{P}}{L} k_1(\lambda) + \frac{\abs{D}}{L} k_2(\lambda)
\]
where
\begin{align}
k_1(\lambda) &= \prob R_1\big( \sin(\lambda + \Phi_1) - \sin(\Phi_1) \big), \;\;\; \text{and} \nonumber\\
k_2(\lambda) &= \prob R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big). \label{eq:k2def}
\end{align}
Since,
\[
\sin(\lambda + \phi) - \sin(\phi) = \lambda \big( \cos(\phi) + \zeta(\lambda) \big),
\]
where $\zeta(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero, then
\begin{align*}
k_1(\lambda) &= \prob R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
&= \lambda \big( \prob R_1\cos(\Phi_1) \; + \; \zeta(\lambda) \prob R_1 \big) \\
&= \lambda \big( 1 \; + \; \zeta(\lambda) K \big),
\end{align*}
since $h_1(0) = \prob R_1\cos(\Phi_1) = 1$ and where we put $K = \prob R_1$, a finite constant.  Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e. $\zeta(\hat{\lambda}_L) = o_P(1)$.  So, 
\[
k_1(\hat{\lambda}_L) = \hat{\lambda}_L \big( 1 \; + \; \zeta(\hat{\lambda}_L) K \big) = \hat{\lambda}_L \big( 1  + o_P(1)\big).
\]

We now analyse the behaviour of $k_2(\lambda)$.  Firsty consider when, $0 \leq \lambda < \tfrac{\pi}{M}$, then for each integer $k$ and when $\phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda )$,

@@

&= \lambda\big(\cos\fracpart{\phi} + \zeta_k(\lambda)\big).

@@

where $\zeta_l(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.  Alternatively when $\phi \in [ \tfrac{(2k+1)\pi}{M} - \lambda,\tfrac{(2k+1)\pi}{M} )$,

@@

@@

Observe that $v_k(\tfrac{(2k+1)\pi}{M}) = 2\sin(\tfrac{\pi}{M})$.  Now, for all $\phi \in \reals$ we have

@@

and since $v_k(\lambda,\phi) = 0$ for $\phi$ in the interval $[0,2\pi)$ when $k \notin \{0,1,\dots,M-1\}$ the infinite sum above can be truncated so that
